---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# klamathFishData

<!-- badges: start -->
<!-- badges: end -->

The **`klamathFishData`**  package is an R data package developed by FlowWest to provide access to curated environmental and biological datasets on fish in the Klamath River Basin. It is designed to support research, modeling, and restoration planning.

## Installation

You can install the development version of klamathFishData from [GitHub](https://github.com/Klamath-SDM/klamathFishData/tree/main) with:

``` r
# install.packages("devtools")
devtools::install_github("Klamath-SDM/klamathFishData")
```
## Data sources

The data in this package comes from a variety of public sources, including:

  - California Department of Fish and Wildlife (CDFW)
  - US Geological Survey (USGS)
  - Literature-based habitat models

## Example

If you are interested in spawner abundance of fall run Chinook salmon:

```{r example}
library(klamathFishData)
library(tidyverse)

fisheries_model_estimates |> 
  filter(estimate_type == "abundance", species == "fall chinook", lifestage == "spawner")
```

## What's Included

The package contains cleaned, documented data objects including:

- `fisheries_model_estimates` — abundance and survival estimates from fisheries models (USE ME)
- `data_location_lookup` - latitude and longitude for fisheries data
- `megatable` - Digital version of [CDFW's Megatable](https://wildlife.ca.gov/Conservation/Fishes/Chinook-Salmon/Anadromous-Assessment)
- And more in development...
 
Use `data(package = "klamathFishData")` to view all datasets included.

