---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# klamathFishData

<!-- badges: start -->
<!-- badges: end -->

The **`klamathFishData`**  package is an R data package developed by FlowWest to provide access to curated environmental and biological datasets on fish in the Klamath River Basin. It is designed to support research, modeling, and restoration planning.

## Installation

You can install the development version of klamathFishData from [GitHub](https://github.com/Klamath-SDM/klamathFishData/tree/main) with:

``` r
# install.packages("devtools")
devtools::install_github("Klamath-SDM/klamathFishData")
```
## Data sources

The data in this package comes from a variety of public sources, including:

  - California Department of Fish and Wildlife (CDFW)
  - US Geological Survey (USGS)
  - Literature-based habitat models
  - more...
  

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(klamathFishData)
## basic example code
```

## What's Included

The package contains cleaned, documented data objects including:

- `rst_sites` —   rotary screw traps
- `hatcheries` — hatchery locations and metadata  
- `habitat_data` — modeled habitat availability by stream and species  
- `model_output` — prespawn mortality and abundance estimates from modeling  
- `klamath_cdfw_population_processed` — CDFW salmonid population estimates  
- And more...
 
Use `data(package = "klamathFishData")` to view all datasets included.

